### 2018-06-09 dissemination.publication/640fb4c4b9ba9822/paper_abstract.tex

Internet of things (IoT) devices capture and create various
forms of sensor data such as images and videos. However, such
resource-constrained devices lack the capability
to efficiently process data in a timely and real-time manner.
Therefore, IoT systems strongly rely on a powerful server
(either local or on the cloud) to extract useful information
from data. In addition, during communication with servers,
unprocessed, sensitive, and private data is transmitted
throughout the Internet, a serious vulnerability. What if we
were able to harvest the aggregated computational power
of already existing IoT devices in our system to locally
process this data? In this artifact, we utilize Musical
Chair~\cite{musical-chair}, which enables efficient,
localized, and dynamic real-time recognition by harvesting the
aggregated computational power of these resource-constrained
IoT devices. We apply Musical chair to two well-known image
recognition models, AlexNet and VGG16, and implement them on
a network of Raspberry PIs (up to 11). We compare inference
per second and energy per inference of our systems with Tegra
TX2, an embedded low-power platform with a six-core CPU and
a GPU. We demonstrate that the collaboration of IoT devices,
enabled by Musical Chair, achieves similar real-time
performance without the extra costs of maintaining a server.